Death by Dropdown?<br>Engineer Insightful Shiny Apps<br>with Behavioral Science

### 2025-09-12 conference-materials/2025_posit-conf/draft.qmd
Update layout and styling for presentation slides
- Convert insight box to blockquote for core realization slide
- Reorganize columns in NOTICE section (move chart before code)
- Replace side-by-side layout with r-stack for ANTICIPATE section
- Combine before/after tables into single comparison table for transformation slide
- Adjust column widths and styling for final slides
- Update figure references and timestamps

::: {.insight-box style="margin-top: 300px !important;"}
"Most dashboard failures aren't technical. They're psychological."
:::

@@

::: {.notes}
- 45 seconds
- Think of building dashboards like you would a road trip
- You need to interpret where you're going, notice warning lights...
- [CLICK] "This is the BID framework"
- [CLICK] "Five *iterative* stages to prevent dashboard disasters"
- Like any good road trip, it's not just about the destination...it's about having the right framework to explore and discover along the way
:::

@@

:::: {.columns}
::: {.column width="55%"}
```{r}
#| code-line-numbers: "|2-6|8"
# Minimal input → Maximum insight
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Users report confusion and telemetry shows low engagement"
)

summary(notice_stage)
```
:::

::: {.column width="45%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

ggplot(filters, aes(x = 1, y = filter, fill = usage)) +
  geom_tile(color = "white", size = 1) +
  geom_text(
    aes(label = paste0(usage, "%")),
    size = 5, fontface = "bold",
    color = ifelse(filters$usage > 40, "white", "black")
  ) +
  scale_fill_gradient2(
    low = "#fee5d9", mid = "#fc9272", high = "#a50f15",
    midpoint = 40, guide = "none"
  ) +
  labs(
    title = "Filter Usage Reality",
    subtitle = "14 of 18 filters basically ignored",
    x = "", y = ""
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16, color = "#666")
  )
```
:::
::::

@@

:::: {.columns}
::: {.column width="40%"}
### Same Data, Different Story

```{r}
anticipate_stage <- bid_anticipate(previous_stage = notice_stage)

# Auto-generates bias mitigations
summary(anticipate_stage)
```
:::

::: {.column width="60%"}
![Marketing: 🎉](img/ui-gain-frame.png){style="margin-bottom: 20px;"}

![Finance: 😱](img/ui-loss-frame.png)
:::
::::

@@

:::: {.columns}
::: {.column width="50%"}
### Before

@@

```{r}
#| echo: false

metrics_before <- data.frame(
  metric = c("Filters", "Success Rate", "Time to Insight", "Return Rate"),
  value = c("18", "8%", "258s", "12%")
)

knitr::kable(metrics_before, col.names = c("Metric", "Value"),
            align = c("l", "r"))
```
:::

::: {.column width="50%"}
### After

![](img/ui-after.png)

```{r}
#| echo: false

metrics_after <- data.frame(
  metric = c("Filters", "Success Rate", "Time to Insight", "Return Rate"),
  value = c("3 (+15 advanced)", "85%", "45s", "67%")
)

knitr::kable(metrics_after, col.names = c("Metric", "Value"),
            align = c("l", "r"))
```
:::
::::

::: {.fragment .text-xl style="text-align: center; margin-top: 30px;"}
**Same data. Same goals. 10x better outcomes.**
:::

@@

# Enhanced metrics visualization
metrics <- data.frame(
  Metric = factor(c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate"),
                 levels = c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate")),
  Before = c(258, 35, 32, 12),
  After = c(45, 85, 78, 67),
  Industry = c("All", "All", "All", "All")
)

metrics_long <- metrics %>%
  pivot_longer(cols = c(Before, After), names_to = "State", values_to = "Value")

# Calculate improvements
metrics_long$Improvement <- ifelse(metrics_long$State == "After",
                                  round((metrics_long$Value / metrics$Before - 1) * 100),
                                  NA)

ggplot(metrics_long, aes(x = Metric, y = Value, fill = State)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = ifelse(Metric == "Time to Insight",
                               paste0(Value, "s"),
                               paste0(Value, "%"))),
           position = position_dodge(width = 0.7),
           vjust = -0.5, size = 8, fontface = "bold") +
  geom_text(data = subset(metrics_long, State == "After" & !is.na(Improvement)),
           aes(label = paste0("+", abs(Improvement), "%"), y = Value + 10),
           position = position_dodge(width = 0.7),
           vjust = -0.5, size = 6, color = "#27ae60", fontface = "italic") +
  scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
  labs(title = "BID Framework Impact: 12 Enterprise Deployments",
       subtitle = "Average improvements across healthcare, finance, and retail sectors",
       y = "") +
  theme_minimal(base_size = 20) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 24),
    legend.title = element_blank(),
    plot.title = element_text(size = 32, face = "bold"),
    plot.subtitle = element_text(size = 24),
    axis.text = element_text(size = 20),
    axis.title = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  ylim(0, max(metrics_long$Value) * 1.2)
```

@@

```{r}
#| eval: false

@@

# Get your next steps
result$next_steps
```

::: {.fragment}
### Resources:
- 🔬 **BID Framework**: [github.com/jrwinget/bid-framework](https://github.com/jrwinget/bid-framework)
- 📚 **{bidux} Docs**: [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
- 💬 **Community**: [github.com/jrwinget/bidux/discussions](https://github.com/jrwinget/bidux/discussions)
- 🤝 **Let's Connect!**
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-bluesky" viewBox="0 0 16 16">
  <path d="M3.468 1.948C5.303 3.325 7.276 6.118 8 7.616c.725-1.498 2.698-4.29 4.532-5.668C13.855.955 16 .186 16 2.632c0 .489-.28 4.105-.444 4.692-.572 2.04-2.653 2.561-4.504 2.246 3.236.551 4.06 2.375 2.281 4.2-3.376 3.464-4.852-.87-5.23-1.98-.07-.204-.103-.3-.103-.218 0-.081-.033.014-.102.218-.379 1.11-1.855 5.444-5.231 1.98-1.778-1.825-.955-3.65 2.28-4.2-1.85.315-3.932-.205-4.503-2.246C.28 6.737 0 3.12 0 2.632 0 .186 2.145.955 3.468 1.948"/></svg> [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-linkedin" viewBox="0 0 16 16">
  <path d="M0 1.146C0 .513.526 0 1.175 0h13.65C15.474 0 16 .513 16 1.146v13.708c0 .633-.526 1.146-1.175 1.146H1.175C.526 16 0 15.487 0 14.854zm4.943 12.248V6.169H2.542v7.225zm-1.2-8.212c.837 0 1.358-.554 1.358-1.248-.015-.709-.52-1.248-1.342-1.248S2.4 3.226 2.4 3.934c0 .694.521 1.248 1.327 1.248zm4.908 8.212V9.359c0-.216.016-.432.08-.586.173-.431.568-.878 1.232-.878.869 0 1.216.662 1.216 1.634v3.865h2.401V9.25c0-2.22-1.184-3.252-2.764-3.252-1.274 0-1.845.7-2.165 1.193v.025h-.016l.016-.025V6.169h-2.4c.03.678 0 7.225 0 7.225z"/></svg> [@jrwinget](https://www.linkedin.com/in/jrwinget/)
  - 📧 [contact@jrwinget.com](mailto:contact@jrwinget.com)
:::

@@

:::: {.columns .center}
::: {.column width="30%"}
![](img/hex-bidux.png)
:::

::: {.column width="40%" .text-display}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="30%" .fragment}
*Special thanks to the posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::
::::

>>>>

<!-- ::: {.insight-box style="margin-top: 300px !important;"}
"Most dashboard failures aren't technical. They're psychological."
::: -->

> "Most dashboard failures aren't technical. They're psychological."

@@

::: {.notes}
- 45 seconds
- Think of building dashboards like you would a road trip
- reference grateful dead fans....people are a long strange trip joke here (related to journey theme; it's a brief point to implicitly reference the VM road image on the slide, and sets them up for understanding the hex logo imagery by the end of the talk)
- You need to interpret where you're going, notice warning lights...
- [CLICK] "This is the BID framework"
- [CLICK] "Five *iterative* stages to prevent dashboard disasters"
- Like any good road trip, it's not just about the destination...it's about having the right framework to explore and discover along the way
:::

@@

:::: {.columns}
::: {.column width="45%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

ggplot(filters, aes(x = 1, y = filter, fill = usage)) +
  geom_tile(color = "white", size = 1) +
  geom_text(
    aes(label = paste0(usage, "%")),
    size = 5, fontface = "bold",
    color = ifelse(filters$usage > 40, "white", "black")
  ) +
  scale_fill_gradient2(
    low = "#fee5d9", mid = "#fc9272", high = "#a50f15",
    midpoint = 40, guide = "none"
  ) +
  labs(
    title = "Filter Usage Reality",
    subtitle = "14 of 18 filters basically ignored",
    x = "", y = ""
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16, color = "#666")
  )
```
:::

::: {.column width="55%"}
```{r}
#| code-line-numbers: "|2-6|8"
# Minimal input → Maximum insight
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Users report confusion and telemetry shows low engagement"
)

summary(notice_stage)
```
:::
::::

@@

:::: {.r-stack}
```{r}
anticipate_stage <- bid_anticipate(previous_stage = notice_stage)

# Auto-generates bias mitigations
summary(anticipate_stage)
```

![](img/ui-gain-frame.png){.fragment}

![](img/ui-loss-frame.png){.fragment}

::: {.text-box-overlay-dark .fragment}
**Same Data, Different Story**
:::
::::

@@

:::: {.r-stack}

@@

![](img/ui-after.png){.fragment}

::: {.text-box-overlay-dark .fragment}
| Metric          | Before     | After              |
|---------------- |:---------: |:------------------:|
| Filters         | 18         | 3 (+15 advanced)   |
| Success Rate    | 8%         | 85%                |
| Time to Insight | 258s       | 45s                |
| Return Rate     | 12%        | 67%                |

::: {.fragment .text-xl style="margin-top: 30px;"}
**Same data. Same goals. 10x better outcomes.**
:::
:::
::::

@@

# Enhanced metrics visualization
metrics <- data.frame(
  Metric = factor(
    c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate"),
    levels = c(
      "Time to Insight",
      "Task Completion",
      "User Satisfaction",
      "Return Rate"
    )
  ),
  Before = c(258, 35, 32, 12),
  After = c(45, 85, 78, 67),
  Industry = c("All", "All", "All", "All")
)

metrics_long <- metrics |>
  pivot_longer(cols = c(Before, After), names_to = "State", values_to = "Value")

# Calculate improvements
metrics_long$Improvement <- ifelse(
  metrics_long$State == "After",
  round((metrics_long$Value / metrics$Before - 1) * 100),
  NA
)

ggplot(metrics_long, aes(x = Metric, y = Value, fill = State)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(
    aes(
      label = ifelse(
        Metric == "Time to Insight",
        paste0(Value, "s"),
        paste0(Value, "%")
      )
    ),
    position = position_dodge(width = 0.7),
    vjust = -0.5, size = 8, fontface = "bold") +
  geom_text(
    data = subset(metrics_long, State == "After" & !is.na(Improvement)),
    aes(label = paste0("+", abs(Improvement), "%"), y = Value + 10),
    position = position_dodge(width = 0.7),
    vjust = -0.5, size = 6, color = "#27ae60", fontface = "italic") +
  scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
  labs(
    title = "BID Framework Impact: 12 Enterprise Deployments",
    subtitle = "Average improvements across healthcare, finance, and retail sectors",
    y = ""
  ) +
  theme_minimal(base_size = 20) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 24),
    legend.title = element_blank(),
    plot.title = element_text(size = 32, face = "bold"),
    plot.subtitle = element_text(size = 24),
    axis.text = element_text(size = 20),
    axis.title = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  ylim(0, max(metrics_long$Value) * 1.2)
```

@@

:::: {.columns}
::: {.column width="65%"}
```{r}
#| eval: false

@@

# Get your next steps
result$next_steps
```
:::

::: {.column width="35%" .fragment}
### Resources:
- 🔬 **BID Framework**: [github.com/jrwinget/bid-framework](https://github.com/jrwinget/bid-framework)
- 📚 **{bidux} Docs**: [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
- 💬 **Community**: [github.com/jrwinget/bidux/discussions](https://github.com/jrwinget/bidux/discussions)
- 🤝 **Let's Connect!**
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-bluesky" viewBox="0 0 16 16">
  <path d="M3.468 1.948C5.303 3.325 7.276 6.118 8 7.616c.725-1.498 2.698-4.29 4.532-5.668C13.855.955 16 .186 16 2.632c0 .489-.28 4.105-.444 4.692-.572 2.04-2.653 2.561-4.504 2.246 3.236.551 4.06 2.375 2.281 4.2-3.376 3.464-4.852-.87-5.23-1.98-.07-.204-.103-.3-.103-.218 0-.081-.033.014-.102.218-.379 1.11-1.855 5.444-5.231 1.98-1.778-1.825-.955-3.65 2.28-4.2-1.85.315-3.932-.205-4.503-2.246C.28 6.737 0 3.12 0 2.632 0 .186 2.145.955 3.468 1.948"/></svg> [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-linkedin" viewBox="0 0 16 16">
  <path d="M0 1.146C0 .513.526 0 1.175 0h13.65C15.474 0 16 .513 16 1.146v13.708c0 .633-.526 1.146-1.175 1.146H1.175C.526 16 0 15.487 0 14.854zm4.943 12.248V6.169H2.542v7.225zm-1.2-8.212c.837 0 1.358-.554 1.358-1.248-.015-.709-.52-1.248-1.342-1.248S2.4 3.226 2.4 3.934c0 .694.521 1.248 1.327 1.248zm4.908 8.212V9.359c0-.216.016-.432.08-.586.173-.431.568-.878 1.232-.878.869 0 1.216.662 1.216 1.634v3.865h2.401V9.25c0-2.22-1.184-3.252-2.764-3.252-1.274 0-1.845.7-2.165 1.193v.025h-.016l.016-.025V6.169h-2.4c.03.678 0 7.225 0 7.225z"/></svg> [@jrwinget](https://www.linkedin.com/in/jrwinget/)
  - 📧 [contact@jrwinget.com](mailto:contact@jrwinget.com)
:::
::::

@@

:::: {.columns .center}
::: {.column width="20%"}
![](img/hex-bidux.png){style="box-shadow: none !important; border: none !important; margin-top: 50px;"}
:::

::: {.column width="40%" .text-display style="margin-top: 200px;"}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="20%" .fragment}
*Special thanks to posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::
::::
